---
title: "Use Elastic Net To Model the Kaggle Housing Data"
output: html_notebook
---
# Prepare the data

## Load the libraries 
```{r,message=TRUE,echo=FALSE,warning=TRUE,include=FALSE}
source(file.path('.','library.R'))
```

## Read the data.
The new object will be called t.
```{r,message=TRUE,echo=FALSE,warning=TRUE,include=FALSE}
source(file.path('.','read_kaggle_housing_data.R'))
```
## Include custom functions
```{r,include=FALSE}
source(file.path('.','functions.R'))
```

## Separate the data set into a testing and training data set.
https://sebastiansauer.github.io/train-test/
```{r}
ex <- modelr::resample_partition(t,c(test= 0.3, train = 0.7))
names(ex)
typeof(ex)
model <- lm(SalePriceLog ~ LotArea, data = ex$train)
```
```{r}
model %>% tidy
rmse(model,ex$train)
rmse(model,ex$test)
```
```{r}
rsquare(model,ex$train)
rsquare(model,ex$test)
```
```{r}
mae(model,ex$train)
mae(model,ex$test)
```
```{r}
qae(model,ex$train)
qae(model,ex$test)
```
## Identify the numeric fields
There are three type of fields, nominal character, interval number and ordinal character that 
could be used as interval or nominal character.
```{r}
# Create a list of the fields to be considered
do_not_use <- c('SalePrice','Id')
response   <- 'SalePriceLog'
t_names    <- names(t)[! names(t) %in% c(do_not_use,response)] 
t_names
```
Find character fields
```{r}
char_names <- t %>% magrittr::extract(t_names) %>% find_fields('character')
char_names
```

```{r}
# First, the fields that could be interval or character
nc_names   <- t %>% magrittr::extract(t_names) %>% numeric_character
print(nc_names)
```
```{r}
# Next, the fields that are most likely interval
interval_names   <- t %>% magrittr::extract(t_names) %>% find_fields(c('integer','double'))
print(interval_names)
```
```{r}
t_interval <- t[c('SalePriceLog',interval_names)] %>% filter(complete.cases(.))
head(t_interval)
t_features <- as.matrix(t_interval[fields])

features <- MatrixModels::model.Matrix(~.,t_interval[fields])[,-1]
length(features)
features
```


```{r}
glmnet(x=features,y=t_interval$SalePriceLog)
```

```{r}
MatrixModels::model.Matrix(~.,t[char_names[c(1,2)]])[,-1][1:10,]
```



```{r}
my_formula <- paste0('SalePriceLog ~ ',paste(char_names[c(1,2,4,5,6,7,8,9)],collapse = '+ ', sep=" " ))
my_formula
mm<-model_matrix(t,as.formula(my_formula))
names(mm)
```

```{r}
my_model<-glmnetUtils::glmnet(SalePriceLog ~. ,t[c(char_names[c(1,2,4,5,6,7,8,9)],"SalePriceLog")])
summary(my_model)
glance(my_model)
tidy(my_model)
names(mm)
typeof(mm)
names(my_model)
my_model$lambda
my_model$a0
my_model$beta
plot(my_model)
(my_model$lambda
predict(my_model,s=my_model$lambda[],t)
coef(my_model,s=0)
broom::tidy(my_model)
tmp_coeffs <- coef(my_model, s = "lambda.min")
```
It is known that the ridge penalty shrinks the coefficients of correlated predictors towards each other while the
lasso tends to pick one of them and discard the others. The elastic-net penalty mixes these two; 
```{r}
data(QuickStartExample)
```
```{r}
typeof(x)
histogram(as.vector(x))
```

```{r}
fit <- glmnet(x,y)
```
```{r}
plot(fit)
```
```{r}
library(glmnet)
t_train <- mm %>% 
  filter(row_number() %in% index)
t_test <- mm %>% 
  filter(!row_number() %in% index)
         
glmnet(t_train,y_train,alpha=1)
```



```{r}
interval_names[interval_names %in% nc_names]
```


```{r}
t_train_Y <- t_train %>% select(selected_fields) %>% complete.cases %>% select(SalePriceLog) %>% as.matrix
t_train_Y <- t_train %>% select(selected_fields) %>% complete.cases %>% select(SalePriceLog) %>% as.matrix
t_train_X <- t_train %>% select(numeric_fields) %>% as.matrix
class(t_train_Y)
```

Standardize the numeric fields

```{r}
names(t_train_X)
as.matrix(t_train_X[numeric_fields])
```
```{r}
t_train_X
```


```{r}
fit.lasso <- glmnet(t_train_X,t_train_Y,family="gaussian",alpha=1)
```

