---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
```{r,message=FALSE}
source(file.path('.','library.R'))
```
### Read in the Data
Create a function to read in the data.

```{r}
read_data <- function(name){   
  #' Load a data set from the data directory using a string lookup
  #'
  #' @param name A text string that will match the partial name of a data file in the data directory
  #' @examples
  #'d <- read_data('train')
  my_path <- file.path('.','data')    
  my_dir  <- list.files(my_path)   
  my_file <- grep(name, my_dir,value=TRUE)   
  d <- read_csv(file.path(my_path,my_file))  
}

```


An indicator field for the presence of a second floor is created.
```{r}
t <- read_data('train') %>%
  mutate(SalePriceLog = log(SalePrice)) %>%
  mutate(secondFloor  = `2ndFlrSF` > 0) %>%
  mutate(X1stFlrSF    = `1stFlrSF` ) %>%
  mutate(X2ndFlrSF    = `2ndFlrSF` ) %>%
  mutate(X3SsnPorch   = `3SsnPorch`)  %>%
  select(-`1stFlrSF`, -`2ndFlrSF`, `3SsnPorch` )
```

The columns contain data and each column has a data class.  There are qualitative and quantitative fields.
```{r}
sapply(t,class)
```
The Kaggle data description page give the descriptions of these fields and they can be added to the data set.
```{r}
var.labels <- c(
    SalePrice    = "The property's sale price in dollars." ,
    MSSubClass    = "The building class",
    MSZoning      = "The general zoning classification",
    LotFrontage   = "Linear feet of street connected to property",
    LotArea       = "Lot size in square feet",
    Street        = "Type of road access",
    Alley         = "Type of alley access",
    LotShape      = "General shape of property",
    LandContour   = "Flatness of the property",
    Utilities     = "Type of utilities available",
    LotConfig     = "Lot configuration",
    LandSlope     = "Slope of property",
    Neighborhood  = "Physical locations within Ames city limits",
    Condition1    = "Proximity to main road or railroad",
    Condition2    = "Proximity to main road or railroad (if a second is present)",
    BldgType      = "Type of dwelling",
    HouseStyle    = "Style of dwelling",
    OverallQual   = "Overall material and finish quality",
    OverallCond   = "Overall condition rating",
    YearBuilt     = "Original construction date",
    YearRemodAdd  = "Remodel date",
    RoofStyle     = "Type of roof",
    RoofMatl      = "Roof material",
    Exterior1st   = "Exterior covering on house",
    Exterior2nd   = "Exterior covering on house (if more than one material)",
    MasVnrType    = "Masonry veneer type",
    MasVnrArea    = "Masonry veneer area in square feet",
    ExterQual     = "Exterior material quality",
    ExterCond     = "Present condition of the material on the exterior",
    Foundation    = "Type of foundation",
    BsmtQual      = "Height of the basement",
    BsmtCond      = "General condition of the basement",
    BsmtExposure  = "Walkout or garden level basement walls",
    BsmtFinType1  = "Quality of basement finished area",
    BsmtFinSF1    = "Type 1 finished square feet",
    BsmtFinType2  = "Quality of second finished area (if present)",
    BsmtFinSF2    = "Type 2 finished square feet",
    BsmtUnfSF     = "Unfinished square feet of basement area",
    TotalBsmtSF   = "Total square feet of basement area",
    Heating       = "Type of heating",
    HeatingQC     = "Heating quality and condition",
    CentralAir    = "Central air conditioning",
    Electrical    = "Electrical system",
    X1stFlrSF    = "First Floor square feet",
    X2ndFlrSF    = "Second floor square feet",
    LowQualFinSF  = "Low quality finished square feet (all floors)",
    GrLivArea     = "Above grade (ground) living area square feet",
    BsmtFullBath  = "Basement full bathrooms",
    BsmtHalfBath  = "Basement half bathrooms",
    FullBath      = "Full bathrooms above grade",
    HalfBath      = "Half baths above grade",
    BedroomAbvGr  = "Number of bedrooms above basement level",
    KitchenAbvGr  = "Number of kitchens",
    KitchenQual   = "Kitchen quality",
    TotRmsAbvGrd  = "Total rooms above grade (does not include bathrooms)",
    Functional    = "Home functionality rating",
    Fireplaces    = "Number of fireplaces",
    FireplaceQu   = "Fireplace quality",
    GarageType    = "Garage location",
    GarageYrBlt   = "Year garage was built",
    GarageFinish  = "Interior finish of the garage",
    GarageCars    = "Size of garage in car capacity",
    GarageArea    = "Size of garage in square feet",
    GarageQual    = "Garage quality",
    GarageCond    = "Garage condition",
    PavedDrive    = "Paved driveway",
    WoodDeckSF    = "Wood deck area in square feet",
    OpenPorchSF   = "Open porch area in square feet",
    EnclosedPorch = "Enclosed porch area in square feet",
    X3SsnPorch    = "Three season porch area in square feet",
    ScreenPorch   = "Screen porch area in square feet",
    PoolArea      = "Pool area in square feet",
    PoolQC        = "Pool quality",
    Fence         = "Fence quality",
    MiscFeature   = "Miscellaneous feature not covered in other categories",
    MiscVal       = "$Value of miscellaneous feature",
    MoSold        = "Month Sold",
    YrSold        = "Year Sold",
    SaleType      = "Type of sale",
    SaleCondition = "Condition of sale",
    SalePriceLog  = "Log of Sale Price",
    secondFloor   = "House has Second Floor"
)
```

```{r}
t <- Hmisc::upData(t,labels = var.labels)
Hmisc::label(t)
```
Make sure the labels work.
```{r}
label(t$OverallQual)
label(t$SalePriceLog)
```
```{r}
class(t)
typeof(t)
```

### Plot the data

```{r}
head(label(t))
head(t)
```

```{r}
t %>% ggplot()  + 
  aes(X1stFlrSF,SalePriceLog,color=secondFloor)+
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha=0.35) + 
  ggtitle(paste(label(t$SalePrice),label(t$X1stFlrSF))) +
  #geom_smooth(method='loess',formula = y~x) + 
  #geom_smooth(method='lm',formula = y~x) + 
  xlab(label(t$X1stFlrSF)) +
  ylab(label(t$SalePrice) )
```

## First Model
First we will experiment with the numeric fields. We will leave the category fields aside for now.
```{r}
t_types <- sapply(t,typeof)
t_types
names(t)
keep_fields <- !(names(t) %in% c('SalePrice','Id'))
keep_fields

t_numeric_fields <- t_types[t_types != 'character' & keep_fields ]
t_numeric_fields
t_numeric <- t[names(t_numeric_fields)]
sapply(t_numeric,typeof)

head(t_numeric)
names(t_numeric)

sapply(t_numeric,typeof)
```
### Corrrelation
So now we search for factors that are significantly correlated and order those by strength of correlation.
```{r}
model_num = 1
num_of_canadate_features = 15
sig <- 0.0010
response <- "SalePriceLog"
names(t_numeric)

# Calculate the correlation matrix
res2 <- rcorr(as.matrix(t_numeric))
res2$P
# It is a list of matricies with names "r","n" and "P"
c(typeof(res2),names(res2))

# Identify the significant correlations 
dim(res2$P)

res2_price_p <- res2$P[response,]
res2_price_p
correlation_significant<-res2_price_p <= sig & !is.na(res2_price_p)
```
Subset to only fields significantly correlated with there response
```{r}

# extract the correlations of the response to other fields
res2_price_r <- res2$r[response,]
res2_price_r

# Select the significant correlations
features_significant <- res2_price_r[correlation_significant]
features_significant 

# Sort by correlation
features_significant <- features_significant[order(abs(features_significant),decreasing=TRUE)]
features_significant
names(features_significant)

remove_response <- !names(features_significant) %in% c('SalePriceLog',response)
remove_response
features_significant <- features_significant[remove_response]
features_significant

#redo the t_selected
t_selected <- t_numeric %>% select(c(names(features_significant)[1:num_of_canadate_features],response)) 
names(t_selected)
# Note that the correlation for OverallQual is zero.  It was the field used last time, so the correlation
#should be zero
```

```{r}
features_significant 
names(t_selected) 
```

A few versions of the correlation matrix
```{r}
t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='ellipse',order="AOE")

t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='shade',order="AOE")

t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot.mixed(number.cex = 0.5,order = "AOE",tl.pos='lt')
#t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='circle',method='ellipse')
```

Here is the top feature to try
```{r}
feature_list <-list()
selected<-names(features_significant)[1]
selected 
selected_label<-label(t_numeric[selected])
selected_label
feature_list[[model_num]] <- selected
str(feature_list)
```
Calculate a model using this variable
```{r}
names(t_numeric)
try(t_numeric <- t_numeric %>% select(-.hat,-.sigma,-.cooksd,-.std.resid,-.se.fit,-.fitted,-.resid,-influential),silent=TRUE)
names(t_numeric)
feature_list
my_formula <- paste0('SalePriceLog ~ ',paste(feature_list,collapse = '+ ', sep=" " ))
str(my_formula)
model_1 <- t_numeric %>% lm(my_formula,data=.)
t_numeric<- model_1 %>% augment(type.predict = 'response',type.residuals = 'response',data=t_numeric) %>%
  mutate(model_1_resid = .resid) %>%
  mutate(influential = .cooksd > 4/length(SalePriceLog))
head(t_numeric)
```
Check the fit using the Kaggle criteria
```{r}
rmse(model_1,t_numeric)
```
```{r}
t_numeric[t_numeric$influential,]
```
```{r}
model_1_parameters <- model_1 %>% tidy
print(model_1_parameters)
```
The R^2 is very good.
```{r}
model_1_fit <- model_1 %>% glance
print(model_1_fit)
```

```{r}
t_numeric %>% ggplot()  + 
  aes(.fitted,SalePriceLog,color=influential) +
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha = 0.35) +
  geom_line(aes(.fitted,.fitted)) +
  ggtitle(selected_label) +
  xlab('fitted') +
  ylab("Sale Price")   
```
```{r}
t_numeric %>% ggplot()  + 
  aes(.resid,SalePriceLog,color=influential) +
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha = 0.35) +
  ggtitle(selected_label) +
  xlab('residuals') +
  ylab("Sale Price")   
```
This fitted by residual should show no pattern.
```{r}
t_numeric %>% ggplot()  + 
  aes(.fitted,.resid)+
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha=0.35) + 
  ggtitle(paste("Residuals Plot")) 
```

```{r}
test <- read_data('test') %>%
  mutate(secondFloor  = `2ndFlrSF` > 0) %>%
  mutate(X1stFlrSF    = `1stFlrSF` ) %>%
  mutate(X2ndFlrSF    = `2ndFlrSF` ) %>%
  mutate(X3SsnPorch   = `3SsnPorch`)  %>%
  select(-`1stFlrSF`, -`2ndFlrSF`, `3SsnPorch` )
```

```{r}
p <- test %>%  
  add_predictions(model_1)  %>%
  select(Id,pred) %>%
  mutate(SalePrice = exp(pred)) %>%
  select(Id,SalePrice)
head(p)
write_csv(p,file.path('.','data','model_1.csv'))
```



## Model 2

### Corrrelation
So now we search for factors that are significantly correlated and order those by strength of correlation.
```{r}
names(t_numeric)
try(t_numeric <- t_numeric %>% select(-.hat,-.sigma,-.cooksd,-.std.resid,-.se.fit,-.fitted,-.resid,-influential),silent=TRUE)
names(t_numeric)
```

```{r}
model_num = 2
num_of_canadate_features = 15
sig <- 0.0010
response <- "model_1_resid"
names(t_numeric)

# Calculate the correlation matrix
res2 <- rcorr(as.matrix(t_numeric))
res2$P
# It is a list of matricies with names "r","n" and "P"
c(typeof(res2),names(res2))

# Identify the significant correlations 
dim(res2$P)

res2_price_p <- res2$P[response,]
res2_price_p
correlation_significant<-res2_price_p <= sig & !is.na(res2_price_p)
```
Subset to only fields significantly correlated with there response
```{r}

# extract the correlations of the response to other fields
res2_price_r <- res2$r[response,]
res2_price_r

# Select the significant correlations
features_significant <- res2_price_r[correlation_significant]
features_significant 

# Sort by correlation
features_significant <- features_significant[order(abs(features_significant),decreasing=TRUE)]
features_significant
names(features_significant)

remove_response <- !names(features_significant) %in% c('SalePriceLog',response)
remove_response
features_significant <- features_significant[remove_response]
features_significant

#redo the t_selected
t_selected <- t_numeric %>% select(c(names(features_significant)[1:num_of_canadate_features],response)) 
names(t_selected)
# Note that the correlation for OverallQual is zero.  It was the field used last time, so the correlation
#should be zero
```

```{r}
features_significant 
names(t_selected) 
```

A few versions of the correlation matrix
```{r}
t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='ellipse',order="AOE")

t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='shade',order="AOE")

t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot.mixed(number.cex = 0.5,order = "AOE",tl.pos='lt')
#t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='circle',method='ellipse')
```

Here is the top feature to try
```{r}
#feature_list <-list()
selected<-names(features_significant)[1]
selected 
selected_label<-label(t_numeric[selected])
selected_label
feature_list[[model_num]] <- selected
str(feature_list)
```
Calculate a model using this variable
```{r}

names(t_numeric)
feature_list
my_formula <- paste0('SalePriceLog ~ ',paste(feature_list,collapse = '+ ', sep=" " ))
str(my_formula)
model_2 <- t_numeric %>% lm(my_formula,data=.)
t_numeric<- model_2 %>% augment(type.predict = 'response',type.residuals = 'response',data=t_numeric) %>%
  mutate(model_2_resid = .resid) %>%
  mutate(influential = .cooksd > 4/length(SalePriceLog))
head(t_numeric)
```
Check the fit using the Kaggle criteria
```{r}
rmse(model_2,t_numeric)
```
```{r}
t_numeric[t_numeric$influential,]
```
```{r}
model_2_parameters <- model_2 %>% tidy
print(model_2_parameters)
```
The R^2 is very good.
```{r}
model_2_fit <- model_2 %>% glance
print(model_1_fit)
```

```{r}
t_numeric %>% ggplot()  + 
  aes(.fitted,SalePriceLog,color=influential) +
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha = 0.35) +
  geom_line(aes(.fitted,.fitted)) +
  ggtitle(selected_label) +
  xlab('fitted') +
  ylab("Sale Price")   
```
```{r}
t_numeric %>% ggplot()  + 
  aes(.resid,SalePriceLog,color=influential) +
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha = 0.35) +
  ggtitle(selected_label) +
  xlab('residuals') +
  ylab("Sale Price")   
```
This fitted by residual should show no pattern.
```{r}
t_numeric %>% ggplot()  + 
  aes(.fitted,.resid)+
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha=0.35) + 
  ggtitle(paste("Residuals Plot")) 
```



```{r}
p <- test %>%  
  add_predictions(model_2)  %>%
  select(Id,pred) %>%
  mutate(SalePrice = exp(pred)) %>%
  select(Id,SalePrice)
head(p)
write_csv(p,file.path('.','data','model_2.csv'))
```
## Model 3
```{r}
names(t_numeric)
try(t_numeric <- t_numeric %>% select(-.hat,-.sigma,-.cooksd,-.std.resid,-.se.fit,-.fitted,-.resid,-influential),silent=TRUE)
try(t_numeric <- t_numeric %>% select(-model_1_resid))
names(t_numeric)

```

### Corrrelation
So now we search for factors that are significantly correlated and order those by strength of correlation.
```{r}
names(t_numeric)
model_num = 3
num_of_canadate_features = 15
sig <- 0.0010
response <- "model_2_resid"
names(t_numeric)

# Calculate the correlation matrix
res2 <- rcorr(as.matrix(t_numeric))
res2$P
# It is a list of matricies with names "r","n" and "P"
c(typeof(res2),names(res2))

# Identify the significant correlations 
dim(res2$P)

res2_price_p <- res2$P[response,]
res2_price_p
correlation_significant<-res2_price_p <= sig & !is.na(res2_price_p)
```
Subset to only fields significantly correlated with there response
```{r}

# extract the correlations of the response to other fields
res2_price_r <- res2$r[response,]
res2_price_r

# Select the significant correlations
features_significant <- res2_price_r[correlation_significant]
features_significant 

# Sort by correlation
features_significant <- features_significant[order(abs(features_significant),decreasing=TRUE)]
features_significant
names(features_significant)

remove_response <- !names(features_significant) %in% c('SalePriceLog',response)
remove_response
features_significant <- features_significant[remove_response]
features_significant

#redo the t_selected
t_selected <- t_numeric %>% select(c(names(features_significant)[1:num_of_canadate_features],response)) 
names(t_selected)
# Note that the correlation for OverallQual is zero.  It was the field used last time, so the correlation
#should be zero
```

```{r}
features_significant 
names(t_selected) 
```

A few versions of the correlation matrix
```{r}
t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='ellipse',order="AOE")

t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='shade',order="AOE")

t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot.mixed(number.cex = 0.5,order = "AOE",tl.pos='lt')
#t_selected %>% cor(use='pairwise.complete.obs') %>% corrplot(method='circle',method='ellipse')
```

Here is the top feature to try
```{r}
#feature_list <-list()
selected<-names(features_significant)[1]
selected 
selected_label<-label(t_numeric[selected])
selected_label
feature_list[[model_num]] <- selected
str(feature_list)
```
Calculate a model using this variable
```{r}
names(t_numeric)
feature_list
my_formula <- paste0('SalePriceLog ~ ',paste(feature_list,collapse = '+ ', sep=" " ))
str(my_formula)
model_3 <- t_numeric %>% lm(my_formula,data=.)
t_numeric<- model_3 %>% augment(type.predict = 'response',type.residuals = 'response',data=t_numeric) %>%
  mutate(model_3_resid = .resid) %>%
  mutate(influential = .cooksd > 4/length(SalePriceLog))
head(t_numeric)
```
Check the fit using the Kaggle criteria
```{r}
rmse(model_2,t_numeric)
```
```{r}
t_numeric[t_numeric$influential,]
```
```{r}
model_3_parameters <- model_3 %>% tidy
print(model_3_parameters)
```
The R^2 is very good.
```{r}
model_3_fit <- model_3 %>% glance
print(model_1_fit)
```

```{r}
t_numeric %>% ggplot()  + 
  aes(.fitted,SalePriceLog,color=influential) +
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha = 0.35) +
  geom_line(aes(.fitted,.fitted)) +
  ggtitle(selected_label) +
  xlab('fitted') +
  ylab("Sale Price")   
```
```{r}
t_numeric %>% ggplot()  + 
  aes(.resid,SalePriceLog,color=influential) +
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha = 0.35) +
  ggtitle(selected_label) +
  xlab('residuals') +
  ylab("Sale Price")   
```
This fitted by residual should show no pattern.
```{r}
t_numeric %>% ggplot()  + 
  aes(.fitted,.resid)+
  scale_y_continuous(labels = comma)  + 
  scale_x_continuous(labels = comma) +
  geom_point(alpha=0.35) + 
  ggtitle(paste("Residuals Plot")) 
```



```{r}
p <- test %>%  
  add_predictions(model_3)  %>%
  select(Id,pred) %>%
  mutate(SalePrice = exp(pred)) %>%
  select(Id,SalePrice)
head(p)
write_csv(p,file.path('.','data','model_3.csv'))
```

```{r}
names(t)
```
```{r}
m <- data.frame(model.matrix(Id ~ MSZoning,t))
names(m)
```

```{r}
model <- t %>% lm(SalePriceLog ~ MSZoning,data=.)
model %>% tidy
```
```{r}
model %>% glance
```
```{r}
names(t)
my_corr <- function(feature) {
  f_sub <-substitute(feature)
  my_f <- paste("SalePriceLog ~",f_sub)
  model <- t %>% lm(my_f,data=.)
  model %>% glance
}

my_corr(PavedDrive)

```





